Methods

### 2021-08-12 text/methods.Rmd
methods

## 2. Quantification of functions
\noindent For each transect, we estimated five key process-based functions mediated by fishes: N excretion rate (gN/day/m^2^), P excretion rate (gP/day/m^2^), production of body mass through growth (gC/day/m^2^), herbivory, i.e. ingestion rate of macrophytes (gC/day/m^2^), and piscivory, i.e. ingestion rate of fishes (g/day/m^2^) [@Brandl2019front]. These five functions were estimated in each transect using individual-based bioenergetic models that predict elemental fluxes, including ingestion rate, excretion rates of N and P, and growth rate.
The bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. We quantified the input parameters, including elements of metabolism, growth, and diet and body stoichiometry, for all 1110 species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). We ran an individual bioenergetic model for each combination of species identity, body size, and sea surface temperature (n = 30668) to get the contribution of each individual to each process in each transect and the summed estimates on the fish community level per surface area (m^2^). Each function is thus expressed in dry mass (of C, N, or P) per day per square meter. We note that N excretion, P excretion, and biomass production include contributions of all fishes, whereas herbivory and piscivory are carried out by a subset of the community, with respect to their trophic guild [@Parravicini2020]. To reduce the occurrence of misclassification of herbivores and piscivores, we categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion [@Mouillot2014; @Parravicini2020] (Fig. S5). Both the herbivory and piscivory rates are congruent with the expert opinion trophic guild classifications.
Finally, we estimated multifunction, i.e. one measure that combines all five functions by taking the geometric average.

@@

## 5. Imputation of herbivory and piscivory

## 4. Regression models
The standing stock biomass of communities is inevitably positively related to all functions because of the additive nature of the quantification and general metabolic theory. Furthermore, because of the known relationship between temperature and parameters related to growth and respiration, functions are expected to be positively correlated with temperature. To model the effect of biomass and sea surface temperature (sst), while simultaneously investigating the correlations between the the five functions, we performed a multivariate Bayesian mixed effect regression the log-transformed functions for transect-level observations ($i$) including site and location effects:

$$\begin{bmatrix}Y_{ExN,i} \\ Y_{ExP,i} \\ Y_{Prod,i} \\ Y_{Herb,i} \\ Y_{Pisc,i} \end{bmatrix} \sim MVNormal\left(\begin{bmatrix} \mu_{ExN,i} \\ \mu_{ExP,i} \\ \mu_{Prod,i} \\ \mu_{Herb,i} \\ \mu_{Pisc,i}\end{bmatrix}, S \right),$$
$$S = \begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix} 
R 
\begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix}$$

$$\mu_{\textrm{ExN,i}} = (\beta0_{ExN} +  \delta_{ExN, loc} +  \delta_{ExN, site}) +{\beta1_{ExN}} {x_{log(biomass),i}} + {\beta2_{ExN}} {x_{sst,i}} $$
$$\beta0_{ExN} \sim  student( 3, -3.9, 2.5)\\
\beta0_{ExP} \sim  student( 3, -6.6, 2.5)\\
\beta0_{Prod} \sim  student( 3, -4.3, 2.5)\\
\beta0_{Herb} \sim  student( 3, -2.4, 2.5)\\
\beta0_{Pisc} \sim  student( 3, -4.6, 2.5)\\
R \sim lkj\_corr(1)
$$

\noindent We used non-centered parametrisation for site and location effects and all standard deviations had the following prior: $\sigma \sim student(3, 0, 2.5)$. To further optimize convergence of the model, we sampled from the cholesky decomposed triangular matrix L such that $R = LL^T$. We used flat priors for the model slopes $\beta1, \beta2$.

The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by  using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples  were used to estimate posterior distributions. We used Bayesian R2 to estimate the amount of explained variation of each model. Posterior predictive checks for are provided in Supporting Informationxxxxx.

>>>>

## 2. Quantification of functions
\noindent For each transect, we estimated five key process-based functions mediated by fishes: N excretion rate (gN/day/m^2^), P excretion rate (gP/day/m^2^), production of body mass through growth (gC/day/m^2^), herbivory, i.e. ingestion rate of macrophytes (gC/day/m^2^), and piscivory, i.e. ingestion rate of fishes (g/day/m^2^) [@Brandl2019front]. These five functions were estimated in each transect using individual-based bioenergetic models that predict elemental fluxes, including ingestion rate, excretion rates of N and P, and growth rate.
The bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. We quantified the input parameters, including elements of metabolism, growth, and diet and body stoichiometry, for all 1110 species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). We ran an individual bioenergetic model for each combination of species identity, body size, and sea surface temperature (n = 30668) to get the contribution of each individual to each process in each transect and the summed estimates on the fish community level per surface area (m^2^). Each function is thus expressed in dry mass (of C, N, or P) per day per square meter. We note that N excretion, P excretion, and biomass production include contributions of all fishes, whereas herbivory and piscivory are carried out by a subset of the community, with respect to their trophic guild [@Parravicini2020]. To reduce the occurrence of misclassification of herbivores and piscivores, we categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion [@Mouillot2014; @Parravicini2020] (Fig. S5). Both the herbivory and piscivory rates are congruent with the expert opinion trophic guild classifications.
Finally, we estimated multifunction, i.e. one measure that combines all five functions by taking the geometric average of the five standardized functions. Using the geometric mean rather than the arithmetic mean because functions are dependent on each other and may vary considerably.

@@

## 4. Data analysis

### 4.1 Imputation of herbivory and piscivory
In the underwater visual transect database, 291 transects (3%) did not contained herbivores and 4467 transects (49%) did not contain piscivores yielding false zeros for herbivory and piscivory, respectively. We considered these zeros as missing values because the absence of the observation of a herbivore or piscivore does not prove the actual absence of herbivores and piscivores. In fact, it is highly unlikely that a coral reef fish community contains no herbivores and piscivores at all. To avoid removing all transects with missing values for herbivory or piscivory (n = 4620) from our database when running multivariate analyses, we imputed the missing values by fitting two models using the observed herbivory and piscivory rates and then imputing the missing values by taking the means of posterior predictions. In other words, we used the predictor values corresponding to missing responses to predict herbivory and piscivory rates. As predictor variables, we used all community structure variables described above and sea surface temperature. Additionally, we added random effects for localities and sites.

First, we simultaneously investigated the effects of biomass and sst on all functions and the correlations among functions (independent of biomass and sst). 
The standing stock biomass of communities is inevitably positively related to all functions because of the additive nature of the quantification and metabolic theory [@Brown2004]. Further, because of the known relationship between temperature and parameters related to growth and respiration (see suppl. methods), functions are expected to be affected by temperature. To model the effects of biomass and sst while simultaneously investigating the correlations between the five functions, we fitted a multivariate Bayesian mixed-effect model using transect-level log-transformed functions that included random effects for sites and localities:

$$\begin{bmatrix}y_{ExN,i} \\ y_{ExP,i} \\ y_{Prod,i} \\ y_{Herb,i} \\ y_{Pisc,i} \end{bmatrix} \sim MVNormal\left(\begin{bmatrix} \mu_{ExN,i} \\ \mu_{ExP,i} \\ \mu_{Prod,i} \\ \mu_{Herb,i} \\ \mu_{Pisc,i}\end{bmatrix}, S \right),$$

$$S = \begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix} 
R 
\begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix}$$

$$\mu_{\textrm{ExN,i}} = (\beta0_{ExN} +  \delta_{ExN, loc} +  \delta_{ExN, site}) +{\beta1_{ExP}} {x_{log(biomass),i}} + {\beta2_{ExN}} {x_{sst,i}} \\
\mu_{\textrm{ExP,i}} = (\beta0_{ExP} +  \delta_{ExP, loc} +  \delta_{ExP, site}) +{\beta1_{ExP}} {x_{log(biomass),i}} + {\beta2_{ExP}} {x_{sst,i}} \\
\mu_{\textrm{Prod,i}} = (\beta0_{Prod} +  \delta_{Prod, loc} +  \delta_{Prod, site}) +{\beta1_{Prod}} {x_{log(biomass),i}} + {\beta2_{Prod}} {x_{sst,i}} \\
\mu_{\textrm{Herb,i}} = (\beta0_{Herb} +  \delta_{Herb, loc} +  \delta_{Herb, site}) +{\beta1_{Herb}} {x_{log(biomass),i}} + {\beta2_{Herb}} {x_{sst,i}} \\
\mu_{\textrm{Pisc,i}} = (\beta0_{Pisc} +  \delta_{Pisc, loc} +  \delta_{Pisc, site}) +{\beta1_{Pisc}} {x_{log(biomass),i}} + {\beta2_{Pisc}} {x_{sst,i}} 
$$

\noindent All parameters $\beta$ were given weakly-informative priors ($\beta1 \sim  normal(1, 1), \beta2 \sim  normal(0, 1)$). We applied non-centered parametrisation for site and location effects ($\delta_{loc}, \delta_{site}$) which also included the correlation between site and locality effects across functions. The model thus included a total of three correlation matrices that each had the same prior ($R \sim lkj\_corr(1)$). All standard deviations had a general weakly-informative prior: $\sigma \sim student(3, 0, 2.5)$.

The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by  using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples were used to estimate posterior distributions. The convergence and fit of the models were verified by examining the Rhat, parameter trace plots, and posterior prediction plots. Posterior predictive checks for are provided in Supporting Information xxxx.
